---
title: "R tutorial of scBiG"
output: html_document
---
In this tutorial, we go over how to use basic scBiG functionality in R. If users have done preprocessing with `Seurat` and would like to continue to use scBiG, they can refer to this tutorial. It requires `reticulate` and `sceasy`, please install `reticulate`, `sceasy` and `scbig` first. However, for more involved analyses, we suggest using scBiG from Python.

## Load required packages
```{r packages,message = FALSE,warning = FALSE}
library(Seurat)
library(dplyr)
library(patchwork)
library(R.utils)
library(rhdf5)
library(hdf5r)
library(sceasy)
library(Matrix)
library(igraph) 
library(reticulate)
library(anndata)
```

## Specify Python environment
Here is the virtual environment used during scBig testing. Users can modify it according to actual conditions.
```{r,warning = FALSE}
use_condaenv("base")
```

## Load data
This data could be downloaded from <https://drive.google.com/drive/folders/1JX589QWH7N895cAMAdJR9KDaFEigy1r3>.

Here is the example dataset (`h5` file) where `X` is the count matrix and `Y` is the cell type label. Notably, scBiG is an unsupervised architecture and the cell type label is solely used for visualization and calculation of clustering metrics. 

You can specify the path of your data here:
```{r}
name  <- "Human_pancreatic_islets"
dataPath <- paste0("./", name, ".h5")
raw_count <- h5read(dataPath, "X")
celltype <- h5read(dataPath, "Y")
celltype <- as.character(celltype)
colnames <- colnames(raw_count, do.NULL = FALSE)
rownames <- rownames(raw_count, do.NULL = FALSE)
colnames(raw_count) <- colnames
rownames(raw_count) <- rownames
real.data <- raw_count
```

## Create Seurat objects and data pre-processing
```{r}
options(Seurat.object.assay.version = "v3")
data <- CreateSeuratObject(counts = real.data,min.cells = 3, min.features = 200)
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
# Calculate cs_factor
data@meta.data$cs_factor <- data@meta.data$nCount_RNA / median(data@meta.data$nCount_RNA)
data@meta.data$celltype <- celltype
# Calculate gs_factor
data[["RNA"]]@meta.features$gs_factor <- apply(data[["RNA"]]$data, 1, max)

# Select high variable genes (Optional)
data <- FindVariableFeatures(data, verbose = FALSE)
top2000 <- head(VariableFeatures(data), 2000)
data <- data[top2000]
```

## Convert Seurat objects into AnnData objects
```{r}
adata <- convertFormat(data, from="seurat", to="anndata", main_layer="counts", drop_single_values=FALSE)
adata$X = adata$X$todense()
adata$raw = adata
```

## Train scBiG
In default mode, the function `run_scbig` takes a preprocessed AnnData object as input. The output returns the updated AnnData object with cell embeddings stored in `adata.obsm['feat']` and louvain clustering results stored in `adata.obs['louvain']`.

In addition, function `run_scbig` would print and return all results during each logging training:

- all_loss: loss of training

- ari_k: ARI of KMeans

- ari_l: ARI of Louvain

- nmi_k: NMI of KMeans

- nmi_l: NMI of Louvain
```{r}
scbig <- import("scbig", convert = FALSE)

scbig$setup_seed(as.integer(100))  #setup seed of scBiG
print("Training scBiG...")
result <- scbig$run_scbig(adata, cl_type = 'celltype', return_all = TRUE)
adata <- result[[0]]
record <- result[[1]]
```

## Visualization
```{r}
cell_feat <- py_to_r(adata$obsm$get('feat'))
gene_feat <- py_to_r(adata$varm$get('feat'))
rownames(cell_feat) <- colnames(data)
rownames(gene_feat) <- rownames(data)

data@reductions$scbig <- CreateDimReducObject(embeddings = cell_feat, loadings = gene_feat, key='scbig', assay='RNA')
data <- FindNeighbors(data, reduction = 'scbig', dims = 1:64)
data <- FindClusters(data, resolution = 1, cluster.name = "scbig_clusters")
data <- RunUMAP(data, reduction = "scbig", dims = 1:64, reduction.name = "umap.scbig")

DimPlot(
    data,
    reduction = "umap.scbig",
    group.by = c("celltype", "scbig_clusters"),
    combine = FALSE, label.size = 2
)

```